---
title: "Model Evaluation Audit"
author: "Alicja Gosiewska"
date: "`r Sys.Date()`"
output: html_document
vignette: >
  %\VignetteEngine{knitr::knitr}
  %\VignetteIndexEntry{Model Evaluation Audit}
  %\usepackage[UTF-8]{inputenc}
---


```{r setup, echo = FALSE}
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```


# Model Evaluation Audit for Classification Problem

In this vignette we present plots for classification models evaluation.

```{r}
library(auditor)
library(mlbench)
```

### Dataset

We work on PimaIndianDiabetes dataset.

```{r}
data(PimaIndiansDiabetes)
head(PimaIndiansDiabetes)
```

We transform dependent variable into binary vector.

```{r}
pima <- PimaIndiansDiabetes
pima$diabetes <- ifelse(pima$diabetes == "pos", 1, 0)
```

### Models

We fit 2 models: glm and svm.

```{r}
model_glm <- glm(diabetes~., data = pima, family = binomial)

library(e1071)
model_svm <- svm(diabetes~., data = pima)
```

### Model Audit

First step of auditing is creating `modelAudit` object. It’s an object that can be used to audit a model. It wraps up a model with meta-data.

```{r}
au_glm <- audit(model_glm, data = pima, y = pima$diabetes)
au_svm <- audit(model_svm, data = pima, y = pima$diabetes, label = "svm")
```


### Receiver Operating Characteristic (ROC)

`modelAudit` object can be used for plotting charts.

```{r}
plotROC(au_glm, au_svm)
```

### LIFT Chart

```{r}
plotLIFT(au_glm, au_svm)
```



# Other methods

Other methods and plots are described in vignettes: 

* [Model Residuals audit](https://mi2datalab.github.io/auditor/articles/model_residuals_audit.html)

* [Model Fit audit](https://mi2datalab.github.io/auditor/articles/model_fit_audit.html)

* [Model Performance audit](https://mi2datalab.github.io/auditor/articles/model_performance_audit.html)

* [Observation Influence audit](https://mi2datalab.github.io/auditor/articles/observation_influence_audit.html)


